### 2013-12-27 paper/make_figures10.tex

    \documentclass[]{article}
\usepackage{graphicx}
\usepackage[paperwidth=4in, paperheight=6in, top=0in, bottom=0in, left=0in, right=0in]{geometry}

\thispagestyle{empty}

\begin{document}
\centering

\centering
\includegraphics[width=0.4\textwidth,trim=0in 0.7in 0in 0in, clip]{molec_figures/molecule300_balls}
    \includegraphics[width=0.4\textwidth,trim=0in 0.7in 0in 0in, clip]{molec_figures/molecule300_wire}\\
    (a)

\centering
  \includegraphics[width=0.4\textwidth,trim=0in 0.7in 0in 0in, clip]{molec_figures/molecule600_balls}
    \includegraphics[width=0.4\textwidth,trim=0in 0.7in 0in 0in, clip]{molec_figures/molecule600_wire}\\

(b)

\centering
  \includegraphics[width=0.4\textwidth,trim=0in 0.7in 0in 0in, clip]{molec_figures/molecule900_balls}
    \includegraphics[width=0.4\textwidth,trim=0in 0.7in 0in 0in, clip]{molec_figures/molecule900_wire}\\
(c)

\end{document}