### 2013-12-27 paper/make_figures3.tex

\documentclass[]{article}
\usepackage{graphicx}
\usepackage[paperwidth=6in, paperheight=6in, top=0in, bottom=0in, left=0in, right=0in]{geometry}

\thispagestyle{empty}

\begin{document}
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[height=6cm]{rxn_figures/rxn_manifold1}\llap{\raisebox{4cm}{\includegraphics[height=2.25cm]{rxn_figures/rxn_manifold1_2}}}\\
(a)
\end{minipage}
\begin{minipage}{0.45\textwidth}
\centering
  \includegraphics[height=6cm]{rxn_figures/rxn_manifold2}\llap{\raisebox{4cm}{\includegraphics[height=2.25cm]{rxn_figures/rxn_manifold2_2}}} \\
(b)
\end{minipage}\\
\begin{minipage}{0.45\textwidth}
\centering
  \includegraphics[height=6cm]{rxn_figures/rxn_manifold3}\llap{\raisebox{4cm}{\includegraphics[height=2.25cm]{rxn_figures/rxn_manifold3_2}}} \\
(c)
\end{minipage}
\begin{minipage}{0.45\textwidth}
\centering
  \includegraphics[height=6cm]{rxn_figures/rxn_manifold4}\llap{\raisebox{4cm}{\includegraphics[height=2.25cm]{rxn_figures/rxn_manifold4_2}}} \\
(d)
\end{minipage}
\end{document}