### 2013-12-27 paper/make_figures4.tex

\documentclass[]{article}
\usepackage{graphicx}
\usepackage[paperwidth=6in, paperheight=5in, top=0in, bottom=0in, left=0in, right=0in]{geometry}

\thispagestyle{empty}

\begin{document}
\centering
\begin{minipage}{0.4\textwidth}
\centering
    \includegraphics[width=\textwidth]{rxn_figures/rxn_NLICA1}
(a)
\end{minipage}
\begin{minipage}{0.4\textwidth}
\centering
      \includegraphics[width=\textwidth]{rxn_figures/rxn_NLICA2}
(b)
\end{minipage}\\
\begin{minipage}{0.4\textwidth}
\centering
\includegraphics[width=\textwidth]{rxn_figures/rxn_NLICA_corr1}
(c)
\end{minipage}
\begin{minipage}{0.4\textwidth}
\centering
  \includegraphics[width=\textwidth]{rxn_figures/rxn_NLICA_corr2}
(d)
\end{minipage}
\end{document}