### 2013-12-27 paper/make_figures4b.tex

\documentclass[]{article}
\usepackage{graphicx}
\usepackage[paperwidth=6in, paperheight=2.75in, top=0in, bottom=0in, left=0in, right=0in]{geometry}

\thispagestyle{empty}

\begin{document}
\centering
\begin{minipage}{0.22\textwidth}
\centering
\includegraphics[width=\textwidth]{rxn_figures/rxn_npoints_embed1}
\includegraphics[width=\textwidth]{rxn_figures/rxn_npoints_spectrum1}
(a)
\end{minipage}
\hfill
\begin{minipage}{0.22\textwidth}
\centering
\includegraphics[width=\textwidth]{rxn_figures/rxn_npoints_embed2}\\
\includegraphics[width=\textwidth]{rxn_figures/rxn_npoints_spectrum2}
(b)
\end{minipage}
\hfill
\begin{minipage}{0.22\textwidth}
\centering
\includegraphics[width=\textwidth]{rxn_figures/rxn_npoints_embed4}\\
\includegraphics[width=\textwidth]{rxn_figures/rxn_npoints_spectrum4}
(c)
\end{minipage}
\hfill
\begin{minipage}{0.22\textwidth}
\centering
\includegraphics[width=\textwidth]{rxn_figures/rxn_npoints_embed6}\\
\includegraphics[width=\textwidth]{rxn_figures/rxn_npoints_spectrum6}
(d)
\end{minipage}
\end{document}