### 2013-12-27 paper/make_figures8.tex

\documentclass[]{article}
\usepackage{graphicx}
\usepackage[paperwidth=6in, paperheight=2.5in, top=0in, bottom=0in, left=0in, right=0in]{geometry}

\thispagestyle{empty}

\begin{document}
\centering
\begin{minipage}{0.3\textwidth}
\centering
    \includegraphics[width=\textwidth]{ala_figures_allodd/ala2_embed1}
    (a)
\end{minipage}
\begin{minipage}{0.3\textwidth}
\centering
    \includegraphics[width=\textwidth]{ala_figures_allodd/ala2_embed2}

(b)
\end{minipage}
\begin{minipage}{0.3\textwidth}
\centering
    \includegraphics[width=\textwidth]{ala_figures_allodd/ala2_embed3}

(c)
\end{minipage}
\end{document}