### 2013-12-27 paper/make_figures9.tex

\documentclass[]{article}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage[paperwidth=4.5in, paperheight=7in, top=0in, bottom=0in, left=0in, right=0in]{geometry}

\thispagestyle{empty}

\begin{document}

  \centering
    \foreach \atomnumber in {1,2,3,6,8} {
        \includegraphics[width=0.3\textwidth]{ala_figures_allodd/ala2_recon_corr_\atomnumber_1}
        \includegraphics[width=0.3\textwidth]{ala_figures_allodd/ala2_recon_corr_\atomnumber_2}
        \includegraphics[width=0.3\textwidth]{ala_figures_allodd/ala2_recon_corr_\atomnumber_3}\\
      }
\end{document}