### 2014-07-17 reducible_SDE/paper/paper_draft.tex
edited paper'

\documentclass[1p]{elsarticle}

\usepackage{amsmath}
\usepackage{amsfonts}

\usepackage{subcaption}

\title{Data-driven Reduction of Multiscale Stochastic Dynamical Systems }

\author[PrincetonCBE]{C.J.~Dsilva}
\ead{cdsilva@princeton.edu}

\author[YaleMath]{R.~Talmon}
\ead{ronen.talmon@yale.edu}

\author[YaleMath]{R.R.~Coifman}
\ead{coifman@math.yale.edu}

\author[PrincetonCBE, PrincetonPACM]{I.G.~Kevrekidis\corref{cor1}}
\ead{yannis@princeton.edu}

\address[PrincetonCBE]{Department of Chemical and Biological Engineering, Princton University, Princeton, New Jersey, 08544, USA}
\address[YaleMath]{Department of Mathematics, Yale University, New Haven, Connecticut, 06520, USA}
\address[PrincetonPACM]{Program in Applied and Computational Mathematics, Princton University, Princeton, New Jersey, 08544, USA}

\cortext[cor1]{Corresponding author}

\begin{document}

\begin{abstract}

\begin{itemize}

\item Link between data mining and dynamical systems

\item Extracting or reducing an SDE system to its slow variables using the Mahalanobis graph and NIVs

\item Full mathematical analysis of the slow variable reduction procedure, including necessary conditions for accurate recovery of the slow variables and a framework for empirically setting the algorithm parameters

\end{itemize}

\end{abstract}

\begin{keyword}

\end{keyword}

\maketitle

\section{Introduction}

\section{Theory}

\subsection{Multiscale SDEs}

Setting
\begin{itemize}
\item system of SDEs with separation of timescales
\item want to recover slow variable(s) even when the fast variables are have large variations
\end{itemize}

Consider equations of the form
\begin{equation} \label{eq:fast_slow_SDE}
\begin{aligned}
dx &= adt + dW^1\\
dy &= -\frac{y}{\epsilon} dt + \frac{1}{\sqrt{\epsilon}} dW^2
\end{aligned}
\end{equation}

So $y$ is a fast noise whose equilibrium measure is bounded and $\mathcal{O}(1)$.
%
We assume that the saturation time of fast scale  of the SDE is smaller than the other relevant timescales.

TODO: should we write the SDE in a more general form?

\subsection{Nonlinear Intrinsic Variables}

Background on NIVs 	
\begin{itemize}

\item Want to recover ``intrinsic'' variables that are independent of the measurement.observation function
\item We define intrinsic as variables that have uncoupled noise with unit variance
\end{itemize}

\begin{equation}
dX^i = a(X) dt + dW^i
\end{equation}
and we will let $X = \begin{bmatrix} X^1 \\ X^2 \\ \vdots \\ X^n \end{bmatrix}$

We want to emphasize the link between multi scale SDEs and NIVs: NIVs rescales the variables so that the ``fast'' variables become small.
%
Consider the SDE in \eqref{eq:fast_slow_SDE}, and let $y = \frac{z}{\sqrt{\epsilon}}$. 
%
Then
\begin{equation}
\begin{aligned}
dx &= adt + dW_1\\
dz &= -\frac{z}{\epsilon} dt +  dW_2
\end{aligned}
\end{equation}
%
Therefore, $z$ is a stochastic variable whose diffusion is unity, and so $x$ and $z$ are the ``intrinsic variables'' for the underlying SDE.
%
However, the equilibrium measure of $z$ is $\mathcal{O}(\epsilon)$, and so when $\epsilon \ll 1$, $z \ll x$, and so we will only recover $x$ (the slow variable).

\subsection{Errors and Estimation}

\begin{itemize}
\item Error and estimation analysis of NIVs (remark that applies to more general settings): give conditions/bounds when we can recover the intrinsic slow variable
%
\item Empirical configuration/demonstration of the method, how to use the analysis to set the parameters in the algorithm
\end{itemize}

Given $X$ governed by the above SDE, we can sample ``bursts'' around a point $X_0$ in order to estimate the covariance.

These clouds will be distributed

\begin{equation}
X \sim \mathcal{N}\left( X_0, \delta t I \right)
\end{equation}

We now consider a function $\mathbf{f}: \mathbb{R}^n \mapsto \mathbb{R}^m$.
%
Let $f^i(X)$ denote the $i^{th}$ entry of $\mathbf{f}$, and let $f^i_j(X) = \frac{\partial f^i}{\partial X^j}$.
%
By Taylor expansion around $X=X_0$,
%
\begin{eqnarray}
\hat{C}(X_0) &=& \delta t J_1 J_1^T 
+ \frac{\delta t^2}{2} \left[ J_1 J_3^T + J_3 J_1^T  
+ \frac{3}{2} J_2 J_2^T 
-\frac{1 }{2} J_2 1 1^T J_2^T \right]
+ \mathcal{O} (\delta t^3) %\\
%&=& \delta t J_1 J_1^T 
%+ \frac{\delta t^2}{2} U \Sigma V
%+ \mathcal{O} (\delta t^3)
\end{eqnarray}
where
\begin{equation}
\begin{aligned}
J_1(X_0) &= \begin{bmatrix}
f_1^1 & f_2^1 & \dots & f_n^1 \\
\vdots & \vdots & \cdots & \vdots \\
f_1^m & f_2^m & \dots & f_n^m
\end{bmatrix}\\
%
J_2(X_0) &= \begin{bmatrix}
f_{11}^1 & f_{22}^1 & \dots & f_{nn}^1 \\
\vdots & \vdots & \cdots & \vdots \\
f_{11}^m & f_{22}^m & \dots & f_{nn}^m
\end{bmatrix}\\
%
J_3(X_0) &= \begin{bmatrix}
f_{111}^1 & f_{222}^1 & \dots & f_{nnn}^1 \\
\vdots & \vdots & \cdots & \vdots \\
f_{111}^m & f_{222}^m & \dots & f_{nnn}^m
\end{bmatrix}
\end{aligned}
\end{equation}
%
In this equation, we see what we expect (a linear function of the variance, due to the gradient of $f$) and a second (higher-order) term arising from the curvature and third derivative of $f$.
%
The importance of this second term increases as the curvature increases, as well as when $\delta t$ increases.

Then
\begin{equation}
\hat{C}^{-1}(X_0) \approx \frac{1}{\delta t} \left( J J^T \right)^{-1} - A
\end{equation}
%
where 
\begin{equation}
A = \frac{1}{2} (J_1 J_1^T)^{-1} U \left(\Sigma^{-1} + \frac{\delta t}{2} V \left( J_1 J_1^T \right)^{-1} U \right)^{-1} V \left( J_1 J_1^T \right)^{-1}
\end{equation}
and
\begin{equation}
U \Sigma V = J_1 J_3^T + J_3 J_1^T + \frac{3}{2} J_2 J_2^T -\frac{1 }{2} J_2 1 1^T J_2^T 
\end{equation}
$A$ is a function of the first-, second-, and third-order derivatives of $\mathbf{f}$.

We now approximate the distances $\| X_2 - X_1 \|$.
%
Let $g = f^{-1}: \mathbb{R}^m \mapsto \mathbb{R}^n$, and let $Y_1 = f(X_1)$, and $Y_2 = f(X_2)$.
%
Again, we let $g_j^i = \frac{\partial g^i}{\partial Y^j}$.
%
Then, by Taylor expansion, we obtain
\begin{equation} \label{eq:taylor_expansion}
\begin{aligned}
&\| X_2 - X_1 \|^2 \\
\le&  \left| \frac{1}{2} (Y_2 - Y_1 )^T ((J J^T)^{-1} (Y_1) + (J J^T)^{-1}(Y_2)) (Y_2 - Y_1 ) \right| \\
& + n \left( \left| K_1 K_2 \right| + \left| \frac{ K_2^2}{4} \right|  + \left| \frac{K_1 K_3}{3} \right|  \right) \| Y_2 - Y_1 \| ^4  \\
& + \mathcal{O} (\|Y_1 - Y_2 \|^6 ) 
\end{aligned}
\end{equation}
%
where
%
\begin{equation}
\begin{aligned}
K_1 &= \sup_{i,j,Y} |g_j^i(Y)|\\
K_2 &= \sup_{i,j,k,Y} |g_{jk}^i(Y)|\\
K_3 &= \sup_{i,j,k,l,Y} |g_{jkl}^i(Y)|
\end{aligned}
\end{equation}

Substituting our estimate of the covariance into \eqref{eq:taylor_expansion}, we obtain
\begin{equation}
\begin{aligned}
&\| X_2 - X_1 \|^2 \\
\approx &  \frac{1}{2} (Y_2 - Y_1)^T \left((J J^T)^{-1}(Y_1)  + (J J^T)^{-1}(Y_2) \right) (Y_2 - Y_1) \\
& - \frac{\delta t}{2} (Y_2 - Y_1)^T \left( A(Y_1) + A(Y_2) \right) (Y_2 - Y_1)\\
& + n \left( \left| K_1 K_2 \right| + \left| \frac{ K_2^2}{4} \right|  + \left| \frac{K_1 K_3}{3} \right|  \right) \| Y_2 - Y_1 \| ^4  \\
& + \mathcal{O} (\|Y_1 - Y_2 \|^6 ) 
\end{aligned}
\end{equation}
%
We want the first term in this expression to dominate.

\section{Results}

 Toy problem results (half-moon): illustration of the analysis + extraction of slow variable and ``suppression'' of fast variable depending on $\delta t$ and saturation time of fast

 Define the following two-dimensional system of SDEs
\begin{eqnarray} \label{eq:init_data}
dX(1) &=& 3 dt + dW_1 \\ 
dX(2) &=& \frac{-X(2)}{\epsilon} dt + dW_2 
\end{eqnarray}
where $W_1, W_2$ are independent Brownian motions, and $\epsilon \ll 1$.

Let
\begin{eqnarray}\label{eq:transformed_data}
f^1(X) &=& X(1) + \frac{ X(2)^2}{\epsilon} \\
f^2(X) &=& \frac{X(2)}{\sqrt{\epsilon}}
\end{eqnarray}
Therefore, $X(2)$ is now a ``fast'' variable, and we have further applied a nonlinear transformation to this fast-slow system.

The inverse function $g = f^{-1}$ is
\begin{eqnarray}
g^1(Y) &=& Y(1) - Y(2)^2 \\
g^2(Y) &=& Y(2) \sqrt{\epsilon}
\end{eqnarray}

\begin{figure}[h]
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{data_init}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{data_transformed}
\caption{}
\end{subfigure}
\caption{(a) The original data, simulated from \eqref{eq:init_data}. (b) The data from (a), transformed according to \eqref{eq:transformed_data}. }
\label{fig:initial_data}
\end{figure}

\subsection{NIVs versus DMAPS}

DMAPS gives the fast, NIVs gives the slow

\begin{figure}[h]
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{data_transformed_colored_NIV}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{data_transformed_colored_DMAPS}
\caption{}
\end{subfigure}
\caption{(a) The data from Figure~\ref{fig:initial_data}(b), colored by the first NIV. Note that we recover the slow variable. (b) The data from Figure~\ref{fig:initial_data}(b), colored by the first DMAPS variable. Note that we do {\em not} recover the slow variable.  }
\label{fig:NIV_versus_DMAPS}
\end{figure}

\subsection{Determining Appropriate Parameters}

Illustration of different aspects (error terms) in the analysis

The three relevant quantities for NIVs are 
\begin{itemize}
\item The linearized approximation to the distance,
%
\begin{equation}
(Y_2 - Y_1)^T (JJ^T)^{-1} (Y_2 - Y_1) 
\approx \mathcal{O} \left(  \frac{1}{2} \left( 2 + \epsilon + \sqrt{ 9 + 2 \epsilon + \epsilon^2}\right) \Delta Y^2\right) 
\end{equation}
where $\Delta Y =  \|Y_2 - Y_1\|$

\item The error from the {\em estimation} of the covariance.
%
This is
\begin{equation}
\frac{\delta t}{2} (Y_2 - Y_1)^T A (Y_2 - Y_1) \approx \mathcal{O} \left( \frac{38 \delta t}{\epsilon ^2 + 2 \delta t}\Delta Y^2 \right)
\end{equation}

\item The errors from the {\em truncation} of the Taylor expansion for the Mahalanobis distance, which is 
\begin{equation}
n \left( \left| K_1 K_2 \right| + \left| \frac{ K_2^2}{4} \right|  + \left| \frac{K_1 K_3}{3} \right|  \right) \| Y_2 - Y_1 \| ^4 \approx \mathcal{O} \left( 10 \Delta Y^4  \right) 
\end{equation}

\end{itemize}

\begin{figure}[h]
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{errors_function_dt}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{empirical_totaldist_function_dt}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{errors_function_dy}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{empirical_totaldist_function_dy}
\caption{}
\end{subfigure}
\caption{(a) Analytical estimation of errors as a function of $\delta t$. 
%
(b) Empirical estimation of total distance as a function of $\delta t$.  
%
(c) Analytical estimation of errors as a function of $\Delta Y$. 
%
(d) Empirical estimation of total distance as a function of $\Delta Y$}
\end{figure}

\subsection{Recovery of Fast Variable}

Show empirically what happens as we change the neighborhood (example from notes 1) (recovery of fast variable deep in the spectrum)

\section{Conclusion}

We showed that in certain cases (when we do {\em not} have a simulator where we can change $\delta t$), the data cannot be processed as-is (we cannot find the right kernel scale given a fixed $\delta t$ such that we can accurately recover the slow variable).

If the cloud of samples is too big then we can observe the cloud of clouds (and those clouds can be histograms, Fourier, scattering, etc.) as a way to get smaller clouds.

Richardson extrapolation allows us to get an estimate of a second-order term in the covariance estimation, thereby locally approximating the function using a quadratic form, rather than a linear form,  which can lead to a better/improved/more accurate ``Mahalanobis'' metric.

\end{document}